---
title: "p8105_hw2_rs4338"
author: "Rebecca Shyu"
date: "2024-09-28"
output: github_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

## Problem 0:

* Create a public GitHub repo + local R Project: p8105_hw2_rs4338
* Create a single .Rmd file named p8105_hw2_rs4338.Rmd that renders to github_document
* Create a subdirectory (data) to store the local data files, and use relative paths to access these data files
* Submit a link to your repo via Courseworks
  - https://github.com/rysgpd/p8105_hw2_rs4338

## Problem 1:

* Reading & Cleaning the Dataset
  - Before processing, there were 32 variables and 1,868 rows included in the dataset. It looks like each row is an entrance and/or exit to a station, so there can be many rows for a single station. It provides information about the location (longitude/latitude of both entrance/exit and station) and characteristics (ex: ADA accessible, vending, staffing, etc) of the entrance/exit. There are also up to 11 routes that may stop at the station. 
  - Some data cleaning steps I took were: cleaning names using the `janitor` package,  converted route8 to route11 to `character` when reading in the csv file, selecting the variables outlined in the homework assignment (line, station, name, station latitude / longitude, routes served, entrance type, entry, vending, and ADA compliance) through the `select` function, changed the entry and vending variables from YES/NO to 0/1 from the instructions through the `mutate` function.
  - The dimensions of the resulting table are 19 variables/columns and 1,868 rows/entrances/exits.
  - The data is tidy because the columns are variable names not values, each column has multiple variables, and there is only one table.

```{r prob1_readcsv, message=FALSE}
initial_nyc_transit_df= read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") 

nyc_transit_df = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
                          col_types = cols(
                            Route8 = "c",
                            Route9 = "c",
                            Route10 = "c",
                            Route11 = "c",
                            )
                          ) %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  mutate(
    entry = ifelse(entry == "YES", 1, 0),
    vending = ifelse(vending == "YES", 1, 0)
  )

nyc_transit_df
```

* Answer the following questions:
  - There are 465 distinct stations (using the `distinct` function)
  - There are 84 distinct stations that are ADA-compliant. 
  - The proportion of station entrances/exits w/o vending that allow entrance is 0.3770492.
  - Reformatting the data using the `unite` function, there are 
  
  How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?


```{r probl1_questions, collapse = TRUE}

nyc_transit_df %>% 
  distinct(line, station_name) %>% 
  count()

nyc_transit_df %>% 
  distinct(line, station_name, .keep_all = TRUE) %>% 
  filter(ada == TRUE) %>% 
  count()

no_vending_num = 
  nyc_transit_df %>% 
  filter(
    vending == 0
    ) %>% 
  count()

no_vending_yes_entry =
  nyc_transit_df %>% 
  filter(
    vending == 0,
    entry == 1
  ) %>% 
  count()

no_vending_yes_entry/no_vending_num

    
```

```{r }

```

